---
---
---

**globcci ➔ Independent variable**

**arrival➔ Dependent variable**

**Data: 01/2000 to 12/2019**

-   Train Data: 01/2000 -12/2015

-   Test Data: 01/2016 - 12/2019

**Data Preprocessing (detrended, removed variance):**

-   Lambda=0

```{r}
library(forecast)
library(tseries) 
library(vars)
library(ggplot2)
library(readxl)
library(stats)
library(seasonal)
```

# 1.Load Data

```{r}
df <- read_excel("arrival.xlsx")
df$date <- as.Date(df$date)
df <- subset(df, select = -c(sars,gfc,oil,geopol))
head(df)
```

```{r}
df_ts <- ts(df, start = '2000', frequency = 12)
```

## 2.Seasonally adjusted data

```{r}
#plot all time series
plot(df_ts[,'arrival'], ylab="Tourist Arrival", xlab="Year")
plot(df_ts[,'globcci'], ylab="Global Coincidence Index", xlab="Year")
```

```{r}
fit_mult_arrival <- decompose(df_ts[,'arrival'], type="multiplicative")
plot(fit_mult_arrival)

# fit_mult_oil <- decompose(df_ts[,'oil'], type="multiplicative")
# plot(fit_mult_oil)
# 
# fit_mult_geopol <- decompose(df_ts[,'geopol'], type="multiplicative")
# plot(fit_mult_geopol)

# fit_mult_globcci <- decompose(df_ts[,'globcci'], type="multiplicative")
# plot(fit_mult_globcci)
```

```{r}
a <- seas(x = df_ts[,'arrival'], x11 = "")
autoplot(a)

# o <- seas(x = df_ts[,'oil'], x11 = "")
# autoplot(o)
# 
# gp <- seas(x = df_ts[,'geopol'], x11 = "")
# autoplot(gp)

gc <- seas(x = df_ts[,'globcci'], x11 = "")
autoplot(gc)
```

```{r}
deseasonalized_arrival <- df_ts[,'arrival'] / fit_mult_arrival$seasonal
# deseasonalized_oil <- df_ts[,'oil'] / fit_mult_oil$seasonal
# deseasonalized_geopol <- df_ts[,'geopol'] / fit_mult_geopol$seasonal
```

```{r}
#Warning: Guerrero's method for selecting a Box-Cox parameter (lambda) is given for strictly positive data.
#adding 1 to geopol to remove this error

# rows_with_non_positive_values <- df_ts[deseasonalized_geopol <= 0, ]
# print(rows_with_non_positive_values)
```

```{r}
#df_ts[,'geopol'] <- df_ts[,'geopol'] + 1
# deseasonalized_geopol <- deseasonalized_geopol + 1
```

```{r}
seasonally_adjusted_ts <- cbind(df_ts[,'date'],deseasonalized_arrival,df_ts[,'globcci'])
```

```{r}
# Rename the first and last columns
colnames(seasonally_adjusted_ts)[c(1, ncol(seasonally_adjusted_ts))] <- c("date", "deseasonalized_globcci")
```

## 3.Train-test split

```{r}
train <- window(seasonally_adjusted_ts, start = c(2000, 1), end = c(2015, 12))
test <- window(seasonally_adjusted_ts, start = c(2016, 1), end = c(2019, 12))
```

## 4.Lambda values for each component time series

```{r}
# l_arrival <- BoxCox.lambda(train[,'deseasonalized_arrival'])
# l_arrival
# 
# # l_oil <- BoxCox.lambda(train[,'deseasonalized_oil'])
# # l_oil
# # 
# # l_geopol <- BoxCox.lambda(train[,'deseasonalized_geopol'])
# # l_geopol
# 
# l_globcci <- BoxCox.lambda(train[,'deseasonalized_globcci'])
# l_globcci
```

```{r}
arrival_bc <- BoxCox(train[,"deseasonalized_arrival"], lambda = 0)
# oil_bc <- BoxCox(train[,"deseasonalized_oil"], lambda = l_oil)
# geopol_bc <- BoxCox(train[,"deseasonalized_geopol"], lambda = l_geopol)
globcci_bc <- BoxCox(train[,"deseasonalized_globcci"], lambda = 0)
```

```{r}
# # Correlation calculation
# a <- cbind(train[,'date'],arrival_bc,oil_bc,geopol_bc,globcci_bc)
# cor_matrix <- cor(a[, c("arrival_bc", "oil_bc", "geopol_bc", "globcci_bc")])
# 
# # Print the correlation matrix
# print(cor_matrix)
```

```{r}
#except for geopol all other are non-stationary time series as p-value >0.05
adf.test(arrival_bc)
# adf.test(oil_bc)
# adf.test(geopol_bc)
adf.test(globcci_bc)
```

```{r}
#test data boxcox transformation
arrival_bc_test <- BoxCox(test[,"deseasonalized_arrival"], lambda = 0)
# oil_bc_test <- BoxCox(test[,"deseasonalized_oil"], lambda = l_oil)
# geopol_bc_test <- BoxCox(test[,"deseasonalized_geopol"], lambda = l_geopol)
globcci_bc_test <- BoxCox(test[,"deseasonalized_globcci"], lambda = 0)
```

## 4.VAR()

```{r}
VARselect(cbind(arrival_bc, globcci_bc), lag.max = 10)$selection
```

Explored models of order 3,4 as below.

### 4.1 P=3,4

```{r}
var <- VAR(cbind(arrival_bc, globcci_bc), p = 4, type = "both", season = 12)
#coef(var)
```

```{r}
summary(var)
```

residuals have very small correlation.

```{r}
serial.test(var,lags.pt = 10, type = "PT.asymptotic")
```

**p-value \> 0.05 ==\> accept null hypothesis ==\> no serial correlation in the residuals**

```{r}
acf(resid(var)[, 1], main='arrival')
# acf(resid(var)[, 2], main='oil')
# acf(resid(var)[, 3], main='geopol')
acf(resid(var)[, 2], main='globcci')
```

```{r}
acf(residuals(var))
```

```{r}
#checkresiduals(var)
#Error in ggtsdisplay(residuals, plot.type = "histogram", main = main,  : 
#  ggtsdisplay is only for univariate time series
```

```{r}
pred <- predict(var, n.ahead = length(test[,'deseasonalized_arrival']))

# oil.pred <- ts(pred$fcst$oil_bc[,1], st = c(2016, 1), fr = 12)
# geopol.pred <- ts(pred$fcst$geopol_bc[,1], st = c(2016, 1), fr = 12)
globcci.pred <- ts(pred$fcst$globcci_bc[,1], st = c(2016, 1), fr = 12)
arrival.pred <- ts(pred$fcst$arrival_bc[,1], st = c(2016, 1), fr = 12)
```

```{r}
ts.plot(cbind(window(arrival_bc, start = c(2000)), arrival.pred), lty = 1:2)
# ts.plot(cbind(window(oil_bc, start = c(2000)), oil.pred), lty = 1:2)
# ts.plot(cbind(window(geopol_bc, start = c(2000)), geopol.pred), lty = 1:2)
ts.plot(cbind(window(globcci_bc, start = c(2000)), globcci.pred), lty = 1:2)
```

```{r}
plot <- autoplot(window(arrival_bc, start = c(2000))) + 
  autolayer(arrival_bc_test, series = "Test Data") + 
  autolayer(arrival.pred, series = "Forecast") + 
  xlab("Year") + 
  ylab("Seasonally Adjusted and Box-Cox Transformed Arrival") + 
  ggtitle("VAR Forecast")
```

```{r}
plot
```

```{r}
# generated forecast by model
forecast(var) %>%
  autoplot() + xlab("Year")
```

```{r}
accuracy(arrival.pred, arrival_bc_test)
```

```{r}

residuals_list <- as.data.frame(residuals(var))
p_values <- numeric()

for (i in 1:ncol(residuals_list)) {
  ljung_box_test <- Box.test(residuals_list[, i], lag = 10, type = "Ljung-Box")
  p_values <- c(p_values, ljung_box_test$p.value)
}

# Print or analyze the p-values
print(p_values)
```

```{r}
# aicc <- AIC(var, nobs = length(cbind(arrival_bc, globcci_bc)), type = "AICc")
```

```{r}
# accuracy((BoxCox(test[,'arrival'], lambda = l)), arrival.pred)
# # Calculate model performance metrics 
# mse <- mean((log(test_data_ts[,'arrival']) - arrival.pred)^2)
# rmse <- sqrt(mse)
# mae <- mean(abs(log(test_data_ts[,'arrival']) - arrival.pred))
# 
# # Print the performance metrics
# cat("Mean Squared Error (MSE):", mse, "\n")
# cat("Root Mean Squared Error (RMSE):", rmse, "\n")
# cat("Mean Absolute Error (MAE):", mae, "\n")

# mape <- mean(abs((BoxCox(test[,'arrival'], lambda = l) - arrival.pred) / BoxCox(test[,'arrival'], lambda = l))) * 100
# cat("Mean Absolute Percentage Error (MAPE):", mape, "%\n")

# library(Metrics)
# mape(BoxCox(test[,'arrival'], lambda = l), arrival.pred)
```
